---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Project Prioritization

[![lifecycle](https://img.shields.io/badge/Lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis Build Status](https://img.shields.io/travis/prioritizr/ppr/master.svg?label=Linux%20%26%20Mac%20OSX)](https://travis-ci.org/prioritizr/ppr)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/jeffreyhanson/ppr/master.svg?label=Windows)](https://ci.appveyor.com/project/jeffreyhanson/ppr)
[![Coverage Status](https://codecov.io/github/prioritizr/ppr/coverage.svg?branch=master)](https://codecov.io/github/prioritizr/ppr?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ppr)](https://CRAN.R-project.org/package=ppr)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "man/figures/README-", fig.align = "center")
```

```{r, include = FALSE}
devtools::load_all()
h = 3.5
w = 3.5
ow = "400"
```

**This package is still under development and not ready for use. Please do not use this package yet.**

The _ppr R_ package is decision support tool for prioritizing conservation projects. Prioritizations can be developed by maximizing expected species richness, expected phylogenetic diversity, the number of species that meet persistence targets, or identifying a set of projects that meet persistence targets for minimal cost. Constraints (e.g. lock in specific actions) and feature weights can also be specified to further customize prioritizations. After defining a project prioritization problem, solutions can be obtained using exact algorithms, heuristic algorithms, or using random processes. In particular, it is recommended to install the ['Gurobi' optimizer](https://www.gurobi.com) because it can identify optimal solutions very quickly. Finally, methods are provided for comparing different prioritizations and evaluating their benefits.

## Installation

The latest development version of the _ppr R_ package can be installed using the following code. Please note that you will need install the _ggtree_ package from Bioconductor since it is not available on [The Comprehensive R Archive Network](https://cran.r-project.org/).

```{r, eval = FALSE}
if (!require(devtools))
  install.packages("devtools")
devtools::install_github("prioritizr/ppr")
if (!require(ggtree))
  devtools::install_bioc("ggtree")
```

Please note that you will also need to install the [Gurobi optimization suite](http://www.gurobi.com/) and [_gurobi R_ package](https://www.gurobi.com/documentation/8.1/refman/r_api_overview.html). For instructions on installing these software packages, please refer to [this installation guide](https://cran.r-project.org/web/packages/prioritizr/vignettes/gurobi_installation.html).

## Usage

## Citation

Please use the following citation to cite the _ppr R_ package in publications:

**This package is still under development and not ready for use. Please do not use this package yet.**

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("Hanson JO, Schuster R, Strimas-Mackey M, Bennett J, (",format(Sys.time(), "%Y"),"). ppr: Project Prioritization. R package version ",packageDescription("ppr")$Version,". Available at https://github.com/prioritizr/ppr.\n"))
```
