---
title: "Project Prioritization"
output:
  rmarkdown::html_vignette:
    toc: true
    fig_caption: true
    self_contained: yes
fontsize: 11pt
bibliography: references.bib
csl: reference-style.csl
documentclass: article
vignette: >
  %\VignetteIndexEntry{ppr}
  %\VignetteEngine{knitr::rmarkdown_notangle}
---

```{r, include = FALSE}
# define fig sizes
fw <- 7.0
fh <- 4.5

# disable running vignette code during R CMD check
is_check <- ("CheckExEnv" %in% search()) || any(c("_R_CHECK_TIMINGS_",
             "_R_CHECK_LICENSE_") %in% names(Sys.getenv()))
knitr::opts_chunk$set(fig.align = "center", eval = !is_check)
```

## Overview

The _ppr R_ package is decision support tool for prioritizing conservation projects. Prioritizations can be developed by maximizing expected feature richness, expected phylogenetic diversity, the number of feature that meet persistence targets, or identifying a set of projects that meet persistence targets for minimal cost. Constraints (e.g. lock in specific actions) and feature weights can also be specified to further customize prioritizations. After defining a project prioritization problem, solutions can be obtained using exact algorithms, heuristic algorithms, or using random processes. In particular, it is recommended to install the ['Gurobi' optimizer](https://www.gurobi.com) because it can identify optimal solutions very quickly. Finally, methods are provided for comparing different prioritizations and evaluating their benefits.

## Tutorial

## Citation

Please use the following citation to cite the _ppr R_ package in publications:

```{r, echo = FALSE, results = "asis", comment = ""}
cat(paste0("Hanson JO, Schuster R, Strimas-Mackey M, Bennett J, (",format(Sys.time(), "%Y"),"). ppr: Project Prioritization. R package version ",packageDescription("ppr")$Version,". Available at https://github.com/prioritizr/ppr.\n"))
```

## References
